---
title: "Automatic scoring of the Autism-Spectrum Quotient (AQ)"
author: "Dale Barr"
output: html_document
---

## The data

```{r url, echo = FALSE, message = FALSE, results = 'asis'}
library("tidyverse")
url <- "http://talklab.psy.gla.ac.uk/sm_data.csv"

cat("The anonymized data are available at <", url, ">\n", sep = "")
```

## Respondent information

```{r preprocess, echo = FALSE}

## load and pre-process data
dat <- read_csv(url, skip = 1L, 
                col_types = cols(Participate = col_logical()))

## participant info
pinfo <- dat %>%
  select(Id, Gender, Age, Participate)
```

Complete responses were received from `r nrow(pinfo)` participants. The `r pinfo %>% pull(Participate) %>% sum()` who agreed to participate had the following demographics:

```{r demographics, echo = FALSE}
pinfo %>%
  filter(Participate) %>%
  group_by(Gender) %>%
  summarize(N = n(), `Mean Age` = mean(Age), 
            SD = sd(Age)) %>%
  knitr::kable(digits = 1)
```


## AQ distribution

```{r aq_dist, echo = FALSE}
# score the individual responses
resp <- dat %>%
  select(Id, Q1:Q10) %>%
  gather("question", "score", Q1:Q10) %>% # wide to long
  separate(score, c("qualifier", "agreement"), "\\s") %>%
  ## score questions 1, 7, 8, 10 forward, otherwise backward
  mutate(fwd_format = question %in% c("Q1", "Q7", "Q8", "Q10"),
         score = case_when(fwd_format ~ agreement == "Agree",
                           TRUE ~ agreement == "Disagree"))
# calculate scores for each respondent
scores <- resp %>%
  group_by(Id) %>%
  summarize(totscore = as.integer(sum(score))) %>%
  ungroup() %>%
  filter(!is.na(totscore)) %>% # no NAs
  semi_join(pinfo %>% filter(Participate), "Id")

ggplot(scores, aes(totscore)) + 
  geom_bar(stat = "count") +
  scale_x_discrete(limits = 0:10) +
  coord_cartesian(xlim = c(0, 10)) +
  labs(x = "AQ score")
```

## Reference

Baron-Cohen, S., Wheelwright, S., Skinner, R., Martin, J., & Clubley, E. (2001). The autism-spectrum quotient (AQ): Evidence from asperger syndrome/high-functioning autism, malesand females, scientists and mathematicians. *Journal of autism and developmental disorders*, *31*, 5-17.